Exploring Design Space for Differentially Private Bayesian Inference

### 2019-01-05 ICML2019/main.tex
fix problem in section2

\noindent \textbf{Technical Problem Statement.}
\label{subsec_problem}
We are interested in exploring mechanisms for privately releasing the full posterior
distributions derived in section \ref{sec_background}, as opposed to just sampling from them.
It's worth noticing that the posterior distributions are fully characterized
by their parameters, and the family (Beta, Dirichlet distribution) they belong to. Hence, in case of the
Beta-Binomial model we are interested in releasing a private version of the pair of
parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)$, and
in the case of the Dirichlet-multinomial model we are interested in a private version of
$\valpha'=(\valpha + \Delta \valpha)$. \cite{zhang2016differential} and \cite{xiao2012bayesian}
have already attacked this problem by adding independent Laplacian noise to the
parameters of the posteriors. That is, in the case of the Beta-Binomial system,
the value released would be: $(\tilde\alpha,\tilde\beta)=(\alpha +  \widetilde{\Delta \alpha},\beta + n - \widetilde{\Delta \alpha})$
where $\widetilde{\Delta \alpha}\thicksim \lap{\Delta \alpha}{\frac{2}{\epsilon}}$,
and where $\lap{\mu}{\nu}$ denotes a Laplace random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t. to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent data observations $\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system. Details are introduced in Sec. \ref{sec_mechs}.

However, in previous works, the accuracy of the posterior was measured with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are as defined above.
\todo{This
accuracy metric is meaningless when the results released are
distributions rather than numerical values. In contrast, distribution
metrics such as $f$-divergence, Hellinger distance, etc. come into
mind overtly when we are measuring distance between
distributions. }
{\color{blue}im not sure this is a good way to put it...}
This gives us motivation for exploring the design space
of mechanisms by considering different norms. e.g. a norm over space of measures.
to compute the sensitivity and provide guarantees on the accuracy.

We will use the Hellinger distance $\hellinger(\cdot,\cdot)$, 
Given two probability distributions $\textup{P}$ and $\textup{Q}$,
the Hellinger distance between them is defined as:
\[
\todo{def}
\]
When $\textup{P}=\betad(\alpha_1, \beta_1)$ and $\textup{Q}=\betad(\alpha_2, \beta_2)$ the above reduces to: 
\[
  \sqrt{1 - \frac{\betaf(\frac{\alpha_1 + \alpha_2}{2}, \frac{\beta_1 + \beta_2}{2})}{\sqrt{\betaf(\alpha_1,\beta_1)\betaf(\alpha_2,\beta_2)}}}
\]

@@

  \item The accuracy can be improved by using a smooth bound on Hellinger distance's local sensitivity. As shown in Fig. \ref{fig_sensitivity},
    taking advantage of the gap between the global sensitivity and local sensitivity, we can improve the accuracy by applying an upper bound on local sensitivity instead of using global sensitivity.
\begin{figure}[ht]
\centering
\includegraphics[width=0.35\textwidth]{sensitivity_2d.eps}
\caption{{Sensitivity of $\hellinger$}}
\label{fig_sensitivity}
\end{figure}
\end{itemize}

@@

\subsubsection{Using $\ell_1$ Norm with Improved Sensitivity}
 The $\lapmech$ presented above can be improved in our Bayesian inference model, by adding noise scaled to smaller sensitivity in this improved Laplace mechanism $\ilapmech$.\\
 In terms of two adjacent data sets $\dataobs, \dataobs'$, their posterior distributions by Bayesian inference -- $\bysinfer(\dataobs), \bysinfer(\dataobs')$ -- which parameter differs at most in 2 dimensions even though extended to Dirichlet-multinomial mode, i.e., $\Delta \bysinfer \leq 2$. As a results, it is enough to use sensitivity $1$ in 2 dimensions and $2$ in higher dimensions.\\
 By calibrating noise to this improved sensitivity, $\ilapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$ produces output:
 $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)
$
with $\eta \thicksim \lap{0}{\frac{1}{\epsilon}}$ in Beta-binomial model and
$\dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots, 
\alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0}
$
with $\eta_i \thicksim \lap{0}{\frac{2}{\epsilon}}$ in Dirichlet-multinomial model. Details are shown in Mechanism \ref{mech:ilapmech}.
  \begin{algorithm}
  \caption{$\ilapmech$ in Dirichlet-multinomial model}
  \label{mech:ilapmech}
  \begin{algorithmic}
  \REQUIRE observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\dirichlet{\valpha}$, likelihood: $\multinomial{n}{\vtheta}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)$.
  \STATE \quad {\bf Initialize} $\tilde{\valpha} = (0,\dots )$, $k = |\mathcal{X}|$.
  \STATE \quad {\bf For} $i = 1 \dots (k - 1)$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_k} = \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \ENSURE $\dirichlet{\tilde{\valpha}}$
  \end{algorithmic}
  \end{algorithm}

>>>>

\noindent \textbf{Technical Problem Statement.}
\label{subsec_problem}
We are interested in exploring mechanisms for privately releasing the full posterior
distributions derived in section \ref{sec_background}, as opposed to just sampling from them.
It's worth noticing that the posterior distributions are fully characterized
by their parameters, and the family (Beta, Dirichlet distribution) they belong to. Hence, in case of the
Beta-Binomial model we are interested in releasing a private version of the pair of
parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)$, and
in the case of the Dirichlet-multinomial model we are interested in a private version of
$\valpha'=(\valpha + \Delta \valpha)$. \cite{zhang2016differential} and \cite{xiao2012bayesian}
have already attacked this problem by adding independent Laplacian noise to the
parameters of the posteriors. That is, in the case of the Beta-Binomial system,
the value released would be: $(\tilde\alpha,\tilde\beta)=(\alpha +  \widetilde{\Delta \alpha},\beta + n - \widetilde{\Delta \alpha})$
where $\widetilde{\Delta \alpha}\thicksim \lap{\Delta \alpha}{\frac{2}{\epsilon}}$,
and where $\lap{\mu}{\nu}$ denotes a Laplace random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t. to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent data observations $\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system. Details are introduced in Sec. \ref{sec_mechs}.\\
However, in previous works, the accuracy of the posterior was measured with respect to $\ell_1$ norm. That is, an upper bound was given on
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are as defined above.
\todo{However, distance between
distributions are more naturally to be measured by distribution
metrics such as $f$-divergence, Hellinger distance, etc., than the $\ell_1$ norm shown above.}
{\color{blue} changed a bit, still not sure this is a good way to put it...}
This gives us motivation for exploring the design space
of mechanisms by considering different norms. e.g. a norm over space of measures.
to compute the sensitivity and provide guarantees on the accuracy.\\
We will use the Hellinger distance $\hellinger(\cdot,\cdot)$, 
Given two probability distributions $\textup{P}$ and $\textup{Q}$,
the Hellinger distance between them is defined as:
\[
\hellinger^2(P,Q) = \frac{1}{2} \int \bigg(\frac{dP}{d \lambda} - \frac{dQ}{d \lambda}\bigg)^2 d \lambda
\]
When $\textup{P}=\betad(\alpha_1, \beta_1)$ and $\textup{Q}=\betad(\alpha_2, \beta_2)$ the above reduces to: 
\[
 \hellinger(P,Q) =  \sqrt{1 - \frac{\betaf(\frac{\alpha_1 + \alpha_2}{2}, \frac{\beta_1 + \beta_2}{2})}{\sqrt{\betaf(\alpha_1,\beta_1)\betaf(\alpha_2,\beta_2)}}}
\]

@@

  \item The accuracy can be improved by using a smooth bound on Hellinger distance's local sensitivity. As shown in Fig. \ref{fig_sensitivity},
    taking advantage of the gap between the global sensitivity and local sensitivity, we can improve the accuracy by applying an upper bound on local sensitivity instead of using global sensitivity.
\begin{figure}[ht]
\centering
\includegraphics[width=0.32\textwidth]{sensitivity_2d.eps}
\caption{{The local sensitivity of $\hellinger$}}
\label{fig_sensitivity}
\end{figure}
\end{itemize}

@@

\subsubsection{Using $\ell_1$ Norm with Improved Sensitivity}
 The $\lapmech$ presented above can be improved in our Bayesian inference model, by adding noise scaled to smaller sensitivity in this improved Laplace mechanism $\ilapmech$.\\
 In terms of two adjacent data sets $\dataobs, \dataobs'$, their posterior distributions by Bayesian inference -- $\bysinfer(\dataobs), \bysinfer(\dataobs')$ -- which parameter differs at most in 2 dimensions even though extended to Dirichlet-multinomial mode, i.e., $\Delta \bysinfer \leq 2$. As a results, it is enough to use sensitivity $1$ in 2 dimensions and $2$ in higher dimensions.\\
 By calibrating noise to this improved sensitivity, $\ilapmech(\dataobs,\bysinfer, \mathsf{prior}, \mathsf{likelihood}, \epsilon)$ produces output:
 $\betad(\alpha +  \lfloor{\Delta \alpha + \eta}\rfloor^n_0,\beta + n - \lfloor{\Delta \alpha + \eta}\rfloor^n_0)
$
with $\eta \thicksim \lap{0}{\frac{1}{\epsilon}}$ in Beta-binomial model and
$\dirichlet{\alpha_1 +  \lfloor{\Delta \alpha_1 + \eta_1}\rfloor^n_0, \cdots, 
\alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{\Delta \alpha_i + \eta_i}\rfloor^n_0 \rfloor^n_0}
$
with $\eta_i \thicksim \lap{0}{\frac{2}{\epsilon}}$ in Dirichlet-multinomial model. Details are shown in Mechanism \ref{mech:ilapmech}.
  \begin{algorithm}
  \caption{$\ilapmech$ in Dirichlet-multinomial model}
  \label{mech:ilapmech}
  \begin{algorithmic}
  \REQUIRE observed data set $\dataobs \in \mathcal{X}^n$, inference algorithm: $\bysinfer$, 
  prior: $\dirichlet{\valpha}$, likelihood: $\multinomial{n}{\vtheta}$, $\epsilon$
  \STATE \quad {\bf let} $\dirichlet{\valpha'} = \bysinfer(\multinomial{n}{\vtheta}, \dirichlet{\valpha}, \dataobs)$, $k = |\mathcal{X}|$.
  \STATE \quad {\bf Initialize} a vector $\tilde{\valpha} = (0, \dots, 0 )$ of $k$ dimensions .
  \STATE \quad {\bf For} $i = 1 \dots (k - 1)$:
  \STATE \quad \quad  {\bf let} $\eta \sim \lap{0}{\frac{2}{\epsilon}}$.
  \STATE \quad \quad  $\tilde{\alpha_i}=\alpha_i + \lfloor{(\alpha_i' - \alpha_i) + \eta}\rfloor^n_0$ 
  % \RETURN $a_{k+1}$.
  \STATE \quad $\tilde{\alpha_k} = \alpha_k + \lfloor n - \sum_{i = 1}^{k-1}\lfloor{(\alpha_i' - \alpha_i) + \eta_i}\rfloor^n_0 \rfloor^n_0$
  \ENSURE $\dirichlet{\tilde{\valpha}}$
  \end{algorithmic}
  \end{algorithm}
